---
title: "Choose your own project submission"
author: "Nigel Brown"
date: "10/7/2020"
output: 
  pdf_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
repos <- 'https://cran.rstudio.com/'
if(!require(tidyverse)) install.packages("tidyverse", repos = repos)
if(!require(tidymodels)) install.packages("tidymodels", repos = repos)
if(!require(scales)) install.packages("scales", repos = repos)


# set theme for charts
theme_set(theme_minimal())

# load the dataset lhd
lhd <- read_rds(here::here('data', 'lhd.rds'))
```

# Introduction

The goal of this project is to predict house prices in the Greater London Area.

The dataset for the project is made from data sourced from [HM Land Registry](https://www.gov.uk/government/statistical-data-sets/price-paid-data-downloads) and [FreeMapTools](https://www.freemaptools.com/download-uk-postcode-lat-lng.htm). This data has been processed into a dataset titled lhd.rds and consists of the following features:

1. price 
1. postcode
1. property_type
1. new_build
1. district
1. address
1. transaction_year
1. transaction_month
1. latitude
1. longitude
1. outward_code




# Exploritory Data Analysis

```{r}
dim(lhd)
```

# Methods/Analysis

# Results

# Conclusion


# Disclaimers

Contains HM Land Registry data © Crown copyright and database right 2020. This data is licensed under the [Open Government Licence v3.0](http://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/)


**FreeMapTools**

Contains Ordnance Survey data © Crown copyright and database right 2020.

Contains Royal Mail data © Royal Mail copyright and database right 2020.

Source: Office for National Statistics licensed under the [Open Government Licence v3.0](http://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/)

# Appendix
