---
output: github_document
params:
    className: "CYOP Submission"
    courseName: "HarvardX PH125.9x"
    repoName: "HarvardX"
    repoOwner: "Nigel-Brown"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


```{js,echo=FALSE}
code .sourceCode {
    white-space: pre-wrap;       /* Since CSS 2.1 */
    white-space: -moz-pre-wrap;  /* Mozilla, since 1999 */
    white-space: -pre-wrap;      /* Opera 4-6 */
    white-space: -o-pre-wrap;    /* Opera 7 */
    word-wrap: break-word;       /* Internet Explorer 5.5+ */
    overflow-wrap: break-word;
}
```


```{r settings,echo=FALSE, include=FALSE}
className <- params$className
courseName <-  params$courseName
repo <- params$repoName
owner <- params$repoOwner
packages <- sort(params$packages)
```

# `r courseName`  `r className`

This is a repo for the `r courseName` Chose Your Own Project Submission by `r owner`.

The submission for this project is three files:

1. A report in PDF format
1. A report in Rmd format
1. A script in R format that performs a supervised machine learning task

This files may be found in the directory titled CYOP.
The dataset utilized in this project is located in the data directory.
